---
title: "terraTCGAdata Introduction"
author: "Marcel Ramos"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Obtain Terra TCGA data as MultiAssayExperiment}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    number_sections: yes
    toc: true
---

# GitHub Installation

```{r, eval = FALSE}
if (!require("BiocManager"))
    install.packages("BiocManager")
BiocManager::install("LiNk-NY/terraTCGAdata")
```

# Description

Terra workspaces come pre-packaged with TCGA data. The challenge is to make
use of the Terra data model and represent the data as `MultiAssayExperiment`.

# Default Data Workspace

To get a list of available TCGA workspaces, use the `findTCGAworkspaces()`
function: 

```{r}
findTCGAworkspaces()
```

You can then set a package-wide option with the `terraTCGAworkspace` function
and check the setting with the ``.

```{r,eval=FALSE}
terraTCGAworkspace()
getOption("terraTCGAdata.workspace")
```


# Clinical data resources

```{r}
ct <- getClinicalTable(workspace = "TCGA_COAD_OpenAccess_V1-0_DATA")
ct
names(ct)
```

# Clinical data download

```{r}
column_name <- "clin__bio__nationwidechildrens_org__Level_1__biospecimen__clin"
clin <- getClinical(
    columnName = column_name,
    participants = TRUE,
    workspace = "TCGA_COAD_OpenAccess_V1-0_DATA"
)
clin[, 1:6]
dim(clin)
```

# Assay data resources

```{r}
at <- getAssayTable(workspace = "TCGA_COAD_OpenAccess_V1-0_DATA")
at
names(at)
```

# Intermediate function for obtaining only the data

```{r}
sampleTypesTable(workspace = "TCGA_COAD_OpenAccess_V1-0_DATA")
prot <- getAssayData(
    assayName = "protein_exp__mda_rppa_core__mdanderson_org__Level_3__protein_normalization__data",
    sampleCode = c("01", "10"),
    workspace = "TCGA_COAD_OpenAccess_V1-0_DATA"
)
prot[, 1:5]
```

# MultiAssayExperiment

```{r}
mae <- terraTCGAdata(
    clinicalName = "clin__bio__nationwidechildrens_org__Level_1__biospecimen__clin",
    assays =
      c("protein_exp__mda_rppa_core__mdanderson_org__Level_3__protein_normalization__data",
      "rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes_normalized__data"),
    sampleCode = NULL,
    split = FALSE,
    workspace = "TCGA_COAD_OpenAccess_V1-0_DATA"
)
mae
```

