---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# terraTCGAdata

## cBioPortal data and MultiAssayExperiment

### Overview

The `terraTCGAdata` R package aims to import TCGA datasets, as
[MultiAssayExperiment][1], that are available on the terra platform. The
package provides a set of functions that allow the discovery of
relevant datasets. There is one central function and two helper functions.

1. The main function `terraTCGAdata` allows the creation of the
`MultiAssayExperiment` object from the different indicated resources.
2. The `getClinicalTable` and `getAssayTable` functions allow for the discovery
of datasets within the terra data model. The column names from these
tables can be provided as inputs to the `terraTCGAdata` function.

[1]: http://bioconductor.org/packages/MultiAssayExperiment/

## Quick Start

### Installation

To install from Bioconductor (recommended for most users, this will install the
release or development version corresponding to your version of Bioconductor):

```{r,eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
```

To install from GitHub (for bleeding-edge, generally not necessary because
changes here are also pushed to [bioc-devel][2]):

```{r,eval=FALSE}
if (!require("cBioPortalData", quietly = TRUE))
    BiocManager::install("waldronlab/terraTCGAdata")
```

Note. The above generally requires the installation of [Bioconductor devel][2]
and the most recent R version.

[2]: https://contributions.bioconductor.org/use-devel.html?q=devel#mid-october

In order to make full use of this package, a local installation of GCloud
SDK is required. The common use case for this package lies within the
Terra environment. Please see the Bioconductor tutorials for running RStudio
on Terra.

To load the package:

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(AnVIL)
library(terraTCGAdata)
```

## gcloud sdk installation

A valid GCloud SDK installation is required to use the package. Use the
`gcloud_exists()` function from the `AnVIL` package to identify
whether it is installed in your system.

```{r}
gcloud_exists()
```

You can also use the `gcloud_project` to set a project name by specifying
the project argument:

```{r, eval=gcloud_exists()}
gcloud_project()
```

# Default Data Workspace

To get a list of available TCGA workspaces, use the `findTCGAworkspaces()`
function:

```{r, eval=gcloud_exists()}
findTCGAworkspaces()
```

You can then set a package-wide option with the `terraTCGAworkspace` function
and check the setting with the `getOption('terraTCGAdata.workspace')` option.

```{r,eval=gcloud_exists()}
terraTCGAworkspace()
getOption("terraTCGAdata.workspace")
```

# Clinical data resources

In order to determine what datasets to download, use the `getClinicalTable`
function to list all of the columns that correspond to clinical data
from the different collection centers.

```{r, eval=gcloud_exists()}
ct <- getClinicalTable(workspace = "TCGA_COAD_OpenAccess_V1-0_DATA")
ct
names(ct)
```

# Clinical data download

After picking the column in the `getClinicalTable` output, use the column
name as input to the `getClinical` function to obtain the data:

```{r, eval=gcloud_exists()}
column_name <- "clin__bio__nationwidechildrens_org__Level_1__biospecimen__clin"
clin <- getClinical(
    columnName = column_name,
    participants = TRUE,
    workspace = "TCGA_COAD_OpenAccess_V1-0_DATA"
)
clin[, 1:6]
dim(clin)
```

# Assay data resources

We use the same approach for assay data. We first produce a list of assays
from the `getAssayTable` and then we select one along with any sample
codes of interest.

```{r, eval=gcloud_exists()}
at <- getAssayTable(workspace = "TCGA_COAD_OpenAccess_V1-0_DATA")
at
names(at)
```

# Summary of sample types in the data

You can get a summary table of all the samples in the adata by using the
`sampleTypesTable`:

```{r, eval=gcloud_exists()}
sampleTypesTable(workspace = "TCGA_COAD_OpenAccess_V1-0_DATA")
```

# Intermediate function for obtaining only the data

Note that if you have the package-wide option set, the workspace argument
is not needed in the function call.

```{r, eval=gcloud_exists()}
prot <- getAssayData(
    assayName = "protein_exp__mda_rppa_core__mdanderson_org__Level_3__protein_normalization__data",
    sampleCode = c("01", "10"),
    workspace = "TCGA_COAD_OpenAccess_V1-0_DATA",
    sampleIdx = 1:4
)
head(prot)
```

# MultiAssayExperiment

Finally, once you have collected all the relevant column names,
these can be inputs to the main `terraTCGAdata` function:

```{r, eval=gcloud_exists()}
mae <- terraTCGAdata(
    clinicalName = "clin__bio__nationwidechildrens_org__Level_1__biospecimen__clin",
    assays =
      c("protein_exp__mda_rppa_core__mdanderson_org__Level_3__protein_normalization__data",
      "rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes_normalized__data"),
    sampleCode = NULL,
    split = FALSE,
    sampleIdx = 1:4,
    workspace = "TCGA_COAD_OpenAccess_V1-0_DATA"
)
mae
```

We expect that most `OpenAccess_V1-0` cancer datasets follow this data model.
If you encounter any errors, please provide a minimally reproducible example
at https://github.com/waldronlab/terraTCGAdata.
